Factors affecting reliablity of state-space age-structured assessment models

### 2023-09-01 Project_0/paper/paper.Rmd
update project 0 WP

---
title: Factors affecting reliablity of state-space age-structured assessment models
author: 
  - Timothy J. Miller^1^, Britten, Brooks, Fay, Legault, Stock, et al.
date: ''
fontsize: 12pt
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    includes:
      in_header:
      - options.sty
      - journalshortcuts.tex
    keep_tex: yes
    number_sections: yes
  html_document:
    df_print: paged
  word_document: default
header-includes:
- \usepackage{xspace}
- \usepackage{bm}
- \usepackage{caption,graphics}
- \usepackage{graphicx}
- \usepackage{makecell}
- \renewcommand\figurename{Fig.}
- \captionsetup{labelsep=period, singlelinecheck=false}
- \newcommand{\changesize}[1]{\fontsize{#1pt}{#1pt}\selectfont}
- \renewcommand{\arraystretch}{1.5}
- \renewcommand\theadfont{}
csl: canadian-journal-of-fisheries-and-aquatic-sciences.csl
biblio-style: cjfas2.bst
---

@@

# Introduction

Application of state-space models in fisheries stock assessment and management has expanded dramatically within ICES, Canada and the Northeast US \citep{nielsenberg14,cadigan16,stockmiller21}.

Much is known about the reliability of state-space models that are linear or Gaussian \citep{aeberhardetal18}, but applications in fisheries management are nonlinear and typically include multiple types of observations with varying distributional assumptions. We know relatively little about the statistical reliability of such models. Also, there is a wide range of potential random effects structures in assessment models and we  know little about the ability of information criteria to distinguish among such alternative structures.

But those studies focus primarily on Gaussian linear models.
review literature on reliability of hidden/latent process models. Primarily in other fields.

@@

# Methods

Used the WHAM package  \citep[][commit 77bbd94]{stockmiller21} (Miller and Stock 2020). This packages has also been used to configure operating and estimating models for closed loop simulations evaluating index-based assessment methods \citep{legaultetalinpress} and is used for management of haddock, butterfish, plaice, bluefish in the Northeast US.

We completed a simulation study with a number of operating models that can be categorized based on where random effects are assumed: abundance at age, natural mortality, fleet selectivity, or index catchability. For each operating model assumptions about variance of process errors and observations are required and the values we used were based on a review of the range of estimates from recent applications of WHAM in management of stocks of haddock, butterfish, and American plaice in the NE US.

We simulated 100 data sets for each operating model.
For each simulated data from each operating model we fit a set of estimating models.

Y estimating models fit to each

## Operating models

common to all:   
  ages = 1 to 10+, 
  M
  maturity

marginal standard deviations for random effects are defined in tables of operating models.

### Population

There are 10 age classes: ages 1 to 10+.

Spawning was assumed to occur 1/4 of the way through the year.

Natural mortality rate was assumed 0.2 when it was constant and the mean of the time series process for operating models with M random effects.
maturity a50 = 2.89, slope = 0.88

Weight at age was generated with a LVB growth
\[
L_a = L_{\infty}\left(1 - e^{-k(a - t_0)}\right)
\]
where $t_0 = 0$, $L_\infty = 85$, and $k = 0.3$, and a L-W relationship such that 
\[
W_a = \theta_1 L_a^{\theta_2}
\]
where $\theta_1 = e^{-12.1}$ and $\theta_2 = 3.2$.

We assumed a Beverton-Holt stock recruit function with constant pre-recruit mortality parameters for all operating models. All post-recruit productivity components are constant in the NAA and survey catchability process error operating models. Therefore steepness and unfished recruitment are also constant over the time period for those operating models (Miller and Brooks 2021). We specified unfished recruitment = $R_0 = e^{10}$ and $F_{\text{MSY}} = F_{40} = 0.348$ equated to a steepness of 0.69 and $\alpha=0.60$ and $\beta = 2.4 \times 10^{-5}$ for the
\[
N_{1,y} = \frac{\alpha \text{SSB}_{y-1}}{1 + \beta \text{SSB}_{y-1}} 
\]
Beverton-Holt parameterization.

The magnitude of the overfishing assumptions is based on average estimates of overfishing for NE groundfish stocks from Wiedenmann et al. (20XX). Legault et al. (2023) also used similar approaches to defining fishing mortality histories for operating models.

Currenlty, initial population is configured at the equilibrium distribution fishing at $F = 2.5\times F_{\text{MSY}}$.

Initial population was configured at the equilibrium distribution fishing at either $F = 2.5\times F_{\text{MSY}}$ or $F = F_{\text{MSY}}$ for the two alternative fishing histories. That is for a deterministic model, the age composition would not change over time when the fishing mortality was constant at the respective level.

For operating models with time-varying random effects for M, steepness is not constant, but we used the same alpha and beta parameters as other operating models this equates to a steepness and R0 at the mean of the time series process for M. For operating models with time-varying random effects for fishery selectivity, Fmsy is also not constant however we use the same F history as other operating models which corresponds to Fmsy at the mean selectivity parameters.

### Fleets

We assumed a single fleet operating year round for catch observations with logistic selectivity for the fleet with $a_{50} = 5$ and slope = 1. This selectivity is was used to define $F_{\text{MSY}}$ for the Beverton-Holt stock recruitment parameters above. We assumed a logistic-normal distribution for the age-composition observations for the fleet.

### Fishing histories
All operating models assumed one of two different fishing histories. 
One : Fishing mortality is equal to Fmsy (0.348) for the whole 40 year period.
Two : Fishing mortality is 2.5 times Fmsy for the first 20 years then changes to Fmsy for the last 20 years.

### Indices
Two time series of surveys are assumed and observed in numbers rather than biomass for the entire 40 year period with one occurring in the spring (0.25 way through the year) and one in the fall (0.75 way through the year). Actually we have it currently configured that both occur 0.5 way through the year. Catchability of both surveys are assumed to be 0.1.  We assumed logistic selectivity for both indices with $a_{50} = 5$ and slope = 1. We assumed a logistic-normal distribution for the age-composition observations.

### Observation Uncertainty

@@

### Operating models with random effects on numbers at age

24 operating models, 16 Sel re operating models and 16 q re operating models.
Table of process error assumptions

\begin{landscape}
\begin{table}
\caption{Distinguishing characteristics of the operating models with random effects on survival. Standard deviations (SD) are for log-normal distributed indices and logistic normal distributed age composition observations (fleet and indices). Fishing mortality changes after year 20 (of 40) for fishing histories where fishing mortality is not constant.}
{\footnotesize \input{naa_om_tab}}
\end{table}
\end{landscape}

### Operating models with random effects on natural mortality

16 operating models
Table of process error assumptions

NOTE: inv_trans_rho function in set_M.R is mis-defined. Will affect correlation parameters assigned in operating models?

Steepness and BRPs are not constant when M is time-varying (Miller and Brooks 2021). We uses the a/b parameters for the B-H defined for Fmsy = F40 at the mean M = 0.2 as defined above.

\begin{landscape}
\begin{table}
\caption{Distinguishing characteristics of the operating models with random effects on natural mortality. Standard deviations (SD) are for log-normal distributed indices and logistic normal distributed age composition observations (fleet and indices). Fishing mortality changes after year 20 (of 40) for fishing histories where fishing mortality is not constant. For AR1 process errors, $\sigma$ is defined for the marginal distribution of the processes.}
{\input{M_om_tab}}
\end{table}
\end{landscape}

### Operating models with random effects on fleet selectivity

16 operating models
Table of process error assumptions

BRPs are not constant when fleet selectivity is time-varying. We uses the a/b parameters for the B-H defined for Fmsy = F40 at the mean of the time series model for selectivity which is the same as the constant selectivity defined above.

\begin{landscape}
\begin{table}
\caption{Distinguishing characteristics of the operating models with random effects on selectivity. Standard deviations (SD) are for log-normal distributed indices and logistic normal distributed age composition observations (fleet and indices). Fishing mortality changes after year 20 (of 40) for fishing histories where fishing mortality is not constant. For AR1 process errors, $\sigma$ is defined for the marginal distribution of the processes.}
{\input{Sel_om_tab}}
\end{table}
\end{landscape}

### Operating models with random effects on index catchability

16 operating models
Table of process error assumptions

\begin{landscape}
\begin{table}
\caption{Distinguishing characteristics of the operating models with random effects on catchability. Standard deviations (SD) are for log-normal distributed indices and logistic normal distributed age composition observations (fleet and indices). Fishing mortality changes after year 20 (of 40) for fishing histories where fishing mortality is not constant. For AR1 process errors, $\sigma$ is defined for the marginal distribution of the processes.}
{\input{q_om_tab}}
\end{table}
\end{landscape}

## Estimating models

\begin{table}
\caption{Distinguishing characteristics of the estimating models.}
{\scriptsize \input{em_tab}}
\end{table}

32 estimating models
Table of estimating models

1-20 fit to each NAA RE operating model
5-24 fit to each M RE operating model
5-20,25-28 to each sel RE operating model
5-20, 29-32 to each q RE operating model

@@

Make plot of S-R curve, Fmsy = F40
Initial values for BH parameters are the true values.
Initial values for mean R model = true R0.

@@

Summary statitics

Bias

Mohn's rho

marginal AIC

convergence types: 
1) whether the optimization completed
2) nlminb convergence flag = 0
3) max abs gradient < 1e-6
4) No NAs for SE of fixed effects returned from TMB::sdreport
5) No SE NAs AND no SE > 100

# Results

Do each of these by type of operating model (Naa, M, sel, q)

## Convergence performance

convergence types: 
1) whether the optimization completed
2) nlminb convergence flag = 0
3) max abs gradient < 1e-6
4) No NAs for SE of fixed effects returned from TMB::sdreport
5) No SE NAs AND no SE > 100

@@

### R, R+S operating models

@@

### R+M operating models

@@

### R+Sel operating models

@@

### R+q operating models

@@

## AIC performance for process error structure
\clearpage 
### Numbers at age operating models

## Estimating models include alternative random effects options: NAA, M, sel, q
\clearpage

@@

\begin{table}
\caption{NAA operating models, estimating models all estimate a mean recruitment and M estimated.}
{\input{naa_om_em_R_MF_aic_table}}
\end{table}
\clearpage

\clearpage 
## AIC performance for stock-recruit relationship

### R, R+S operating models

@@

### R+M operating models

@@

### R+Sel operating models

@@

### R+q operating models

@@

\clearpage 
## Bias, Mean Square error

  Certain basic parameters (stock-recruit pars, M, variance parameters)
  SSB, F, R, BRPs

\begin{landscape}
\begin{figure}
\caption{Median relative bias of for SSB for estimating models that estimate mean recruitment and M is fixed at the true value.}\label{naa_om_em_R_MF_relbias_ssb}
\begin{center}
\includegraphics[width = \textwidth]{naa_om_R_MF_relbias_ssb.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Median relative bias of for SSB for estimating models that estimate mean recruitment and M is estimated.}\label{naa_om_em_R_ME_relbias_ssb}
\begin{center}
\includegraphics[width = \textwidth]{naa_om_R_ME_relbias_ssb.png}
\end{center}
\end{figure}
\end{landscape}

>>>>

---
title: Factors affecting reliablity of state-space age-structured assessment models
author: 
  - Timothy J. Miller^1^
  - A. Applegate 
  - G. Britten
  - E. N. Brooks
  - G. Fay
  - A. Hansell
  - C. M. Legault
  - B. Muffley
  - J. Wiedenmann
date: ''
fontsize: 12pt
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    includes:
      in_header:
      - options.sty
      - journalshortcuts.tex
    keep_tex: yes
    number_sections: yes
  html_document:
    df_print: paged
  word_document: default
header-includes:
- \usepackage{xspace}
- \usepackage{bm}
- \usepackage{caption,graphics}
- \usepackage{graphicx}
- \usepackage{makecell}
- \renewcommand\figurename{Fig.}
- \captionsetup{labelsep=period, singlelinecheck=false}
- \newcommand{\changesize}[1]{\fontsize{#1pt}{#1pt}\selectfont}
- \renewcommand{\arraystretch}{1.5}
- \renewcommand\theadfont{}
csl: canadian-journal-of-fisheries-and-aquatic-sciences.csl
biblio-style: cjfas2.bst
---

@@

# Introduction {-}

Application of state-space models in fisheries stock assessment and management has expanded dramatically within ICES, Canada and the Northeast US \citep{nielsenberg14,cadigan16,stockmiller21}. State-space approaches that use random effects to parameterize process errors is considered best practice and a requirement for the next generation of stock assessment models.

Much is known about the reliability of state-space models that are linear or Gaussian \citep{aeberhardetal18}, but applications in fisheries management are nonlinear and typically include multiple types of observations with varying distributional assumptions. Furthermore, there is a wide range of potential random effects structures and model parameters that can be treated as random effects in assessment models. We know relatively little about the statistical reliability of such models or the ability of information criteria to distinguish among such alternative structures.

<!-- But those studies focus primarily on Gaussian linear models.
review literature on reliability of hidden/latent process models. Primarily in other fields.
-->

@@

# Methods {-}

Used the WHAM package  \citep[][commit 77bbd94]{stockmiller21} (Miller and Stock 2020). This packages has also been used to configure operating and estimating models for closed loop simulations evaluating index-based assessment methods \citep{legaultetal23} and is used for management of haddock, butterfish, plaice, bluefish in the Northeast US.

We completed a simulation study with a number of operating models that can be categorized based on where process error random effects are assumed: abundance at age (R, R+S), natural mortality (R+M), fleet selectivity (R+Sel), or index catchability (R+q). For each operating model assumptions about variance of process errors and observations are required and the values we used were based on a review of the range of estimates from recent applications of WHAM in management of stocks of haddock, butterfish, and American plaice in the NE US.

78 total operating models.
We simulated 100 data sets for each operating model.
For each simulated data set we fit a set of 20 estimating models.

## Operating models {-}

### Population {-}

The population model tracks 10 age classes: ages 1 to 10+ and we assume spawning occurs 1/4 of the way through the year. The maturity at age was a logistic curve with a50 = 2.89 and slope = 0.88 and assumed known in all estimation models (Figure \ref{om_maturity}).

\begin{figure}
\caption{The proportion mature at age assumed for the population in all operating and estimating models.}\label{om_maturity}
\begin{center}
\includegraphics[width = \textwidth]{om_maturity.png}
\end{center}
\end{figure}

Weight at age was generated with a LVB growth function
$$
L_a = L_{\infty}\left(1 - e^{-k(a - t_0)}\right)
$$
where $t_0 = 0$, $L_\infty = 85$, and $k = 0.3$, and a L-W relationship such that 
$$
W_a = \theta_1 L_a^{\theta_2}
$$
where $\theta_1 = e^{-12.1}$ and $\theta_2 = 3.2$ (Figure \ref{om_waa}).

\begin{figure}
\caption{The weight at age assumed for the population in all operating and estimating models.}\label{om_waa}
\begin{center}
\includegraphics[width = \textwidth]{om_waa.png}
\end{center}
\end{figure}

We assumed a Beverton-Holt stock recruit function with constant pre-recruit mortality parameters for all operating models. All post-recruit productivity components are constant in the NAA and survey catchability process error operating models. Therefore steepness and unfished recruitment are also constant over the time period for those operating models \citep{millerbrooks21}. We specified unfished recruitment = $R_0 = e^{10}$ and $\Fmsy = \Fspr[40] = 0.348$ equated to a steepness of 0.69 and $\alpha=0.60$ and $\beta = 2.4 \times 10^{-5}$ for the
\[
N_{1,y} = \frac{\alpha \text{SSB}_{y-1}}{1 + \beta \text{SSB}_{y-1}} 
\]
Beverton-Holt parameterization (Figure \ref{om_sr}). For OMs without process errors on natural mortality we assumed the rate was assumed 0.2. For OMs with process errors on natural mortality the mean rate was 0.2.

\begin{figure}
\caption{The Beverton-Holt stock recruit relationship assume for all operating models.}\label{om_sr}
\begin{center}
\includegraphics[width = \textwidth]{om_sr.png}
\end{center}
\end{figure}

Two alternative fishing histories were used for operating models. In the first scenario, the stock experiences overfishing for the first 20 years and fishing at \Fmsy for the last 20 years. In the second scenario, the stock is fished at \Fmsy for the entire time period. The magnitude of the overfishing assumptions is based on average estimates of overfishing for NE groundfish stocks from \citet{wiedenmannetal19}. \Citet{legaultetal23} also used similar approaches to defining fishing mortality histories for operating models.

Initial population was configured at the equilibrium distribution fishing at either $F = 2.5\times \Fmsy$ or $F = \Fmsy$ for the two alternative fishing histories. That is for a deterministic model, the age composition would not change over time when the fishing mortality was constant at the respective level.

For operating models with time-varying random effects for M, steepness is not constant, but we used the same alpha and beta parameters as other operating models this equates to a steepness and R0 at the mean of the time series process for M. For operating models with time-varying random effects for fishery selectivity, \Fmsy is also not constant however we use the same F history as other operating models which corresponds to Fmsy at the mean selectivity parameters.

### Fleets {-}

We assumed a single fleet operating year round for catch observations with logistic selectivity for the fleet with $a_{50} = 5$ and slope = 1 (Figure \ref{om_mean_selectivity}). This selectivity is was used to define \Fmsy for the Beverton-Holt stock recruitment parameters above. We assumed a logistic-normal distribution for the age-composition observations for the fleet.

\begin{figure}
\caption{The selectivity at age assumed for the fishing fleet in operating models without selectivity random effects. The mean selectivity at age across time for the fishing fleet in operating models with selectivity random effects.}\label{om_mean_selectivity}
\begin{center}
\includegraphics[width = \textwidth]{om_mean_selectivity.png}
\end{center}
\end{figure}

### Indices {-}

Two time series of surveys are assumed and observed in numbers rather than biomass for the entire 40 year period with one occurring in the spring (0.25 way through the year) and one in the fall (0.75 way through the year). Actually we have it currently configured that both occur 0.5 way through the year. Catchability of both surveys are assumed to be 0.1.  Like the fishing fleet, we assumed logistic selectivity for both indices with $a_{50} = 5$ and slope = 1 and a logistic-normal distribution for the age-composition observations.

### Observation Uncertainty {-}

@@

### Operating models with random effects on numbers at age {-}

The factorial combinations of fishing histories, observation error assumptions, and marginal variance assumptions for recruitment and survival results in 24 different operating models with random effects on recruitment and survival (Table \ref{naa_om_table}).

\begin{landscape}
\begin{table}
\caption{Distinguishing characteristics of the operating models with random effects on survival. Standard deviations (SD) are for log-normal distributed indices and logistic normal distributed age composition observations (fleet and indices). Fishing mortality changes after year 20 (of 40) for fishing histories where fishing mortality is not constant.}\label{naa_om_table}
{\footnotesize \input{naa_om_tab}}
\end{table}
\end{landscape}

### Operating models with random effects on natural mortality {-}

The factorial combinations of fishing histories, observation error assumptions, and marginal variance and autocorrelation assumptions for natural mortality results in 16 different operating models with random effects on natural mortality (Table \ref{M_om_table}). Internally, WHAM treats natural mortality as a log-transformed parameter 
$$
\log M_y = \mu_M + \epsilon_{M,y}
$$
that is a linear combination of a mean $\mu_M$ and any annual random effects marginally distributed as $\epsilon_{M,y} \sim \text{N}\left(0,\sigma_M^2\right)$. The values for the marginal variance and correlation parameters assumed in the operating models are given in Table \ref{M_om_table}.

\begin{landscape}
\begin{table}
\caption{Distinguishing characteristics of the operating models with random effects on natural mortality. Standard deviations (SD) are for log-normal distributed indices and logistic normal distributed age composition observations (fleet and indices). Fishing mortality changes after year 20 (of 40) for fishing histories where fishing mortality is not constant. For AR1 process errors, $\sigma$ is defined for the marginal distribution of the processes.}\label{M_om_table}
{\input{M_om_tab}}
\end{table}
\end{landscape}

### Operating models with random effects on fleet selectivity {-}

The factorial combinations of fishing histories, observation error assumptions, and marginal variance and autocorrelation assumptions for selectivity parameters results in 16 different operating models with random effects on selectivity (Table \ref{sel_om_table}). Internally, WHAM treats selectivity parameter $s$ as a logit-transformed parameter 
$$
\log\left(\frac{p_{s,y}-l_{s}}{u_{s}-p_{s,y}}\right) = \mu_s + \epsilon_{s,y}
$$
that is a linear combination of a mean $\mu_s$ and any annual random effects marginally distributed as $\epsilon_{s,y} \sim \text{N}\left(0,\sigma_s^2\right)$ where the lower and upper bounds of the parameter ($l_s$ and $u_s$) can be specified by the user. All selectivity parameters are either $a_50$ or slope parameters and we assume bounds of 0 and 10 for all selectivity parameters for all operating and estimating models. The values for the marginal variance and correlation parameters assumed in the operating models are given in Table \ref{sel_om_table}.

\begin{landscape}
\begin{table}
\caption{Distinguishing characteristics of the operating models with random effects on selectivity. Standard deviations (SD) are for log-normal distributed indices and logistic normal distributed age composition observations (fleet and indices). Fishing mortality changes after year 20 (of 40) for fishing histories where fishing mortality is not constant. For AR1 process errors, $\sigma$ is defined for the marginal distribution of the processes.}\label{sel_om_table}
{\input{Sel_om_tab}}
\end{table}
\end{landscape}

### Operating models with random effects on index catchability {-}

The factorial combinations of fishing histories, observation error assumptions, and marginal variance and autocorrelation assumptions for catchability results in 16 different operating models with random effects on catchability (Table \ref{q_om_table}). Internally, WHAM treats catchability for index $i$ as a logit-transformed parameter 
$$
\log\left(\frac{q_{i,y}-l_{i}}{u_{i}-q_{i,y}}\right) = \mu_i + \epsilon_{i,y}
$$
that is a linear combination of a mean $\mu_i$ and any annual random effects marginally distributed as $\epsilon_{i,y} \sim \text{N}\left(0,\sigma_i^2\right)$ where the lower and upper bounds of the catchability ($l_i$ and $u_i$) can be specified by the user. Here we assume bounds of 0 and 1000 for all operating and estimating models. For operating and estimation models with process errors on catchability, the temporal variation is only assumed for the first index. The values for the marginal variance and correlation parameters assumed in the operating models are given in Table \ref{q_om_table}

\begin{landscape}
\begin{table}
\caption{Distinguishing characteristics of the operating models with random effects on catchability. Standard deviations (SD) are for log-normal distributed indices and logistic normal distributed age composition observations (fleet and indices). Fishing mortality changes after year 20 (of 40) for fishing histories where fishing mortality is not constant. For AR1 process errors, $\sigma$ is defined for the marginal distribution of the processes.}\label{q_om_table}
{\input{q_om_tab}}
\end{table}
\end{landscape}

## Estimating models {-}

For each data set simulated from an operating model 20 estimating models were fit. A total of 32 different estimating models were fit across all operating models where the subset of 20 depended on the source of process error (Table \ref{em_table}). The first 20 estimating models in Table \ref{em_table} were fit to simulate data sets from R and R+S operating models. Estimating models 5 to 24 in Table \ref{em_table} were fit to simulate data sets from R+M operating models. Estimating models 5 to 20 and 25-28 in Table \ref{em_table} were fit to simulate data sets from R+Sel operating models. Finally, estimating models 5 to 20 and 29-32 in Table \ref{em_table} were fit to simulate data sets from R+q operating models.

\begin{table}
\caption{Distinguishing characteristics of the estimating models.}\label{em_table}
{\scriptsize \input{em_tab}}
\end{table}

@@

Initial values for BH parameters are the true values.
Initial values for mean R model = true R0.

@@

## Measures of reliability {-}

The first measure of reliability we investigated was frequency of convergence when fitting each estimating model to the simulated data sets. There are various ways to assess convergence of the fit. We summarized 5 alternative categories of convergence.
1. Did the optimization routine (stats::nlminb) complete without error?
2. Did the stats::nlminb convergence flag = 0 indicate successful convergence?
3. Was the maximum absolute value of the gradient of the log-likelihood < $1\times10^{-6}$?
4. Did TMB::sdreport provide non-NA values for all fixed effects standard errors?
5. Did TMB::sdreport provide all standard errors < 100?

The first convergence criterion assesses whether the model crashes. The third is a measure that assesses how flat the likelihood is at the optimized point of the likelihood surface. The fourth and fifth criteria are specific to the hessian-based standard error reporting provided by TMB. the TMB::sdreport function will sometimes return standard error estimates even when the calculated hessian is not invertible (fourth criterion). It will also provide large standard errors for some parameters that are not estimated well or are near bounds on the transformed scale that is of primary interest. For example, variance parameters for random effects can often be estimated near 0 when the fit suggests no variation in the random effects or, equivalently, that the model without random effects is better.

### AIC for model selection {-}

We measured the frequency of correct model selection using marginal AIC. For a given operating model we the set of models that were considered all made the same assumptions on whether or not to estimate M and whether a stock-recruit function is assumed.

We also measured the frequency of correctly selecting models with Beverton-Holt stock recruit function assumed over models without the stock-recruit function (null model). Because the proportions of simulations where the correct model was chosen for a given set of operating model factors implied generally poor performance of AIC, we analyzed the effect of variation in SSB on the probability of lower AIC of the BH models. We fit binomial generalize linear models to the observed AIC choice for each data set as a function of log(SD(SSB)) and operating model.

### Bias {-}

We estimated relative bias 
$$
\widehat {\text{RB}}\left(\theta\right) = \frac{1}{N} \sum^N_{i = 1} \frac{\widehat \theta_i - \theta_i}{\theta_i}
$$
of annual SSB and fully-selected fishing mortality for each estimation model and constructed 95\% confidence intervals for the median relative bias using the binomial distribution approach as in \citet{stockmiller21}. We also estimated bias of stock-recruit parameters for estimation models that assumed the Beverton-Holt stock recruit function.

### Mohn's rho {-}

We estimated Mohn's rho for SSB, fully-selected fishing mortality, and recruitment for each estimation model. We estimated 7 peels for each estimation model. We calculated median 95\% confidence intervals for Mohn's rho using the same methods as that for bias.

# Results {-}

## Convergence performance {-}

@@

### R, R+S operating models {-}

Figures \ref{naa_om_em_R_MF_convergence} to \ref{naa_om_em_BH_ME_convergence}

@@

### R+M operating models {-}

Figures \ref{M_om_em_R_MF_convergence} to \ref{M_om_em_BH_ME_convergence}

@@

### R+Sel operating models {-}

Figures \ref{Sel_om_em_R_MF_convergence} to \ref{Sel_om_em_BH_ME_convergence}

@@

### R+q operating models {-}

Figures \ref{q_om_em_R_MF_convergence} to \ref{q_om_em_BH_ME_convergence}

@@

## AIC performance for process error structure {-}

### R, R+S operating models {-}

Figures \ref{naa_om_proportion_best_aic_R_MF} to \ref{naa_om_proportion_best_aic_SR_ME}

\begin{landscape}
\begin{figure}
\caption{Proportion of simulated data sets from OMs with R and R+S process errors where fitted estimating models had lowest marginal AIC. All estimating models estimate mean recruitment rather than a stock-recruit relationship and and M is fixed at the true value.} \label{naa_om_proportion_best_aic_R_MF}
\begin{center}
\includegraphics[width = \textwidth]{naa_om_proportion_best_aic_R_MF.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Proportion of simulated data sets from OMs with R and R+S process errors where fitted estimating models had lowest marginal AIC. All estimating models estimate mean recruitment rather than a stock-recruit relationship and and M is estimated.} \label{naa_om_proportion_best_aic_R_ME}
\begin{center}
\includegraphics[width = \textwidth]{naa_om_proportion_best_aic_R_ME.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Proportion of simulated data sets from OMs with R and R+S process errors where fitted estimating models had lowest marginal AIC. All estimating models estimate a stock-recruit relationship and and M is fixed at the true value.} \label{naa_om_proportion_best_aic_SR_MF}
\begin{center}
\includegraphics[width = \textwidth]{naa_om_proportion_best_aic_SR_MF.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Proportion of simulated data sets from OMs with R and R+S process errors where fitted estimating models had lowest marginal AIC. All estimating models estimate a stock-recruit relationship and and M is estimated.} \label{naa_om_proportion_best_aic_SR_ME}
\begin{center}
\includegraphics[width = \textwidth]{naa_om_proportion_best_aic_SR_ME.png}
\end{center}
\end{figure}
\end{landscape}

### R+M operating models {-}

Figures \ref{M_om_proportion_best_aic_R_MF} to \ref{M_om_proportion_best_aic_SR_ME}

\begin{landscape}
\begin{figure}
\caption{Proportion of simulated data sets from OMs with R+M process errors where fitted estimating models had lowest marginal AIC. All estimating models estimate mean recruitment rather than a stock-recruit relationship and and M is fixed at the true value.} \label{M_om_proportion_best_aic_R_MF}
\begin{center}
\includegraphics[width = \textwidth]{M_om_proportion_best_aic_R_MF.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Proportion of simulated data sets from OMs with R+M process errors where fitted estimating models had lowest marginal AIC. All estimating models estimate mean recruitment rather than a stock-recruit relationship and and M is estimated.} \label{M_om_proportion_best_aic_R_ME}
\begin{center}
\includegraphics[width = \textwidth]{M_om_proportion_best_aic_R_ME.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Proportion of simulated data sets from OMs with R+M process errors where fitted estimating models had lowest marginal AIC. All estimating models estimate a stock-recruit relationship and and M is fixed at the true value.} \label{M_om_proportion_best_aic_SR_MF}
\begin{center}
\includegraphics[width = \textwidth]{M_om_proportion_best_aic_SR_MF.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Proportion of simulated data sets from OMs with R+M process errors where fitted estimating models had lowest marginal AIC. All estimating models estimate a stock-recruit relationship and and M is estimated.} \label{M_om_proportion_best_aic_SR_ME}
\begin{center}
\includegraphics[width = \textwidth]{M_om_proportion_best_aic_SR_ME.png}
\end{center}
\end{figure}
\end{landscape}

### R+Sel operating models {-}

Figures \ref{Sel_om_proportion_best_aic_R_MF} to \ref{Sel_om_proportion_best_aic_SR_ME}

\begin{landscape}
\begin{figure}
\caption{Proportion of simulated data sets from OMs with R+Sel process errors where fitted estimating models had lowest marginal AIC. All estimating models estimate mean recruitment rather than a stock-recruit relationship and and M is fixed at the true value.} \label{Sel_om_proportion_best_aic_R_MF}
\begin{center}
\includegraphics[width = \textwidth]{Sel_om_proportion_best_aic_R_MF.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Proportion of simulated data sets from OMs with R+Sel process errors where fitted estimating models had lowest marginal AIC. All estimating models estimate mean recruitment rather than a stock-recruit relationship and and M is estimated.} \label{Sel_om_proportion_best_aic_R_ME}
\begin{center}
\includegraphics[width = \textwidth]{Sel_om_proportion_best_aic_R_ME.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Proportion of simulated data sets from OMs with R+Sel process errors where fitted estimating models had lowest marginal AIC. All estimating models estimate a stock-recruit relationship and and M is fixed at the true value.} \label{Sel_om_proportion_best_aic_SR_MF}
\begin{center}
\includegraphics[width = \textwidth]{Sel_om_proportion_best_aic_SR_MF.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Proportion of simulated data sets from OMs with R+Sel process errors where fitted estimating models had lowest marginal AIC. All estimating models estimate a stock-recruit relationship and and M is estimated.} \label{Sel_om_proportion_best_aic_SR_ME}
\begin{center}
\includegraphics[width = \textwidth]{Sel_om_proportion_best_aic_SR_ME.png}
\end{center}
\end{figure}
\end{landscape}

### R+q operating models {-}

Figures \ref{q_om_proportion_best_aic_R_MF} to \ref{q_om_proportion_best_aic_SR_ME}

\begin{landscape}
\begin{figure}
\caption{Proportion of simulated data sets from OMs with R+q process errors where fitted estimating models had lowest marginal AIC. All estimating models estimate mean recruitment rather than a stock-recruit relationship and and M is fixed at the true value.} \label{q_om_proportion_best_aic_R_MF}
\begin{center}
\includegraphics[width = \textwidth]{q_om_proportion_best_aic_R_MF.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Proportion of simulated data sets from OMs with R+q process errors where fitted estimating models had lowest marginal AIC. All estimating models estimate mean recruitment rather than a stock-recruit relationship and and M is estimated.} \label{q_om_proportion_best_aic_R_ME}
\begin{center}
\includegraphics[width = \textwidth]{q_om_proportion_best_aic_R_ME.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Proportion of simulated data sets from OMs with R+q process errors where fitted estimating models had lowest marginal AIC. All estimating models estimate a stock-recruit relationship and and M is fixed at the true value.} \label{q_om_proportion_best_aic_SR_MF}
\begin{center}
\includegraphics[width = \textwidth]{q_om_proportion_best_aic_SR_MF.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Proportion of simulated data sets from OMs with R+q process errors where fitted estimating models had lowest marginal AIC. All estimating models estimate a stock-recruit relationship and and M is estimated.} \label{q_om_proportion_best_aic_SR_ME}
\begin{center}
\includegraphics[width = \textwidth]{q_om_proportion_best_aic_SR_ME.png}
\end{center}
\end{figure}
\end{landscape}

<!--
## Estimating models include alternative random effects options: NAA, M, sel, q {-}
\clearpage

@@

\begin{table}
\caption{NAA operating models, estimating models all estimate a mean recruitment and M estimated.}
{\input{naa_om_em_R_MF_aic_table}}
\end{table}
\clearpage 
-->

\clearpage 
## AIC performance for stock-recruit relationship {-}

Tables of frequencies where AIC chooses correct stock-recruit assumption imply poor performance of AIC, but figures showing effect of variation in SSB demonstrate that large variation in SSB is needed to reduce Type II errors for marginal AIC.

### R, R+S operating models {-}

@@

### R+M operating models {-}

@@

### R+Sel operating models {-}

@@

### R+q operating models {-}

@@

\clearpage 
## Bias {-}

### R, R+S OMs {-}

\begin{landscape}
\begin{figure}
\caption{Median relative bias of for SSB for estimating models fitted to data sets simulated with R and R+S process errors.  Estimation models do not assume a stock-recruit function and M is fixed at the true value.}\label{naa_om_em_R_MF_relbias_ssb}
\begin{center}
\includegraphics[width = \textwidth]{naa_om_R_MF_relbias_ssb.png}
\end{center}
\end{figure}
\end{landscape}

\begin{landscape}
\begin{figure}
\caption{Median relative bias of for SSB for estimating models fitted to data sets simulated with R and R+S process errors. Estimation models do not assume a stock-recruit relationship and estimate M.}\label{naa_om_em_R_ME_relbias_ssb}
\begin{center}
\includegraphics[width = \textwidth]{naa_om_R_ME_relbias_ssb.png}
\end{center}
\end{figure}
\end{landscape}
